\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2019-01-14 redaction/submission_JoEG_DHP.tex
unknown
je ne sais pas pourquoi il me demande de commiter, je n'ai rien fait depuis tout à l'heure

Trade costs are central to international economic analysis. In particular, they are known to be a major obstacle to international economic integration and international trade flows. Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies, etc.), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}. Many argue that the latter have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies. However, besides several empirical evidence suggesting that transport costs remain large and deserve attention\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs represent a 21\% markup over production costs. Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows.}, distance as an incompressible cost is still a very important determinant of trade flows\footnote{See \citet{Disdier_Head08}}. In addition, progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden. The latter is a key issue, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from regional liberalization.\smallskip

International transport costs accordingly stand at the heart of this paper. We provide a thorough empirical investigation of the pattern of international transport costs over time based on US imports flows from 1974 to 2013. Using a similar dataset ending in 2004, \cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e. \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects (i.e. changes over time in the baskets of imported goods and/or the origin countries) have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we dig deeper into this issue. As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs. Rather, the evolution of transport costs \textit{per se} constitutes the main driving force behind the overall transport costs decrease, in particular for air transport. In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs. In this respect, we find that pure transport costs declined even less since the 1970s than argued by \cite{hummels2007}.\medskip

>>>>

Trade costs are a major obstacle to international trade flows.
They take many forms: transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies, etc.), policy costs (tariffs  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}.
Their importance vary enormously between different countries and different sectors.
They are still the subject of controversy.
One would think that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies.
It is certainly cheaper in constant dollars to transport one ton of goods now than fifty years ago.
Yet, compared to the value of shipments, several empirical evidence suggesting that transport costs remain large and deserve attention
\footnote{\citet{anderson_wincoop_jel} find that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs represent a 21\% markup over production costs. Further, the sizable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) testifies to the impact of transport costs on trade flows. Distance as an incompressible cost is still a very important determinant of trade flows\footnote{See \citet{Disdier_Head08}}}.
Progress in transportation techniques (which reduces the constant-dollar cost of transport) should be compared with technical progress in production (which reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden.
This an important debate, since the answers to several important (both academic and policy) economic questions depend in large part on how much distance-based costs affects trade, such as, among others, the dynamics of agglomeration, the profitability of outsourcing or the gains from regional liberalization.
This debate is all the more complex as the transport cost burden does not depend only on the size of transport costs, but also on whether they are a proportion of the value of transported goods (and hence do not distort relative prices) or are not related in a linear way to the value of transported goods (and hence do distort relative prices), as would a pure per-ton transport cost schedule .
\smallskip

International transport costs accordingly stand at the heart of this paper.
We provide a thorough empirical investigation of the pattern of international transport costs over time based on US imports flows from 1974 to 2013.
Using a similar dataset ending in 2004, \cite{hummels2007} finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e. \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects (i.e. changes over time in the baskets of imported goods and/or the origin countries) have partly offset the pure transport costs decrease, thereby attenuating the overall downward trend in transport costs.
\footnote{Throughout the paper, we refer to ``pure'' or ``\textit{per se}'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.}
In this paper, we dig deeper into this issue. As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs.
Rather, the evolution of transport costs \textit{per se} constitutes the main driving force behind the overall transport costs decrease, in particular for air transport. In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs.
In this respect, we find that pure transport costs declined even less since the 1970s than argued by \cite{hummels2007}.\medskip
